I can see clearly now: reinterpreting statistical significance

### 2019-01-07 rebuttal.md

Reviewer: 1
===========

_This perspective paper proposes that we change the language we use to report significance of hypothesis tests. They suggest that a simple change in the language will help clarify the meaning and also help highlight the importance of effect sizes and confidence intervals. Overall, it’s well-written and is a good contribution to this important subject area. We should be open to ideas that experiment and potentially improve how we use NHST._

_I agree that a downside of hypothesis testing is often (wrongly) seen as an end in itself, rather than the scientific knowledge that is gained from the test. However, I am not convinced that your suggestion will solve this. Here are my thoughts:_

Thanks for your perspective. We are also not convinced our suggestion will solve this major problem, to put it mildly. But we hope it will help (and think that in small ways it already has). We respond to your specific points below.

@@

_My major concern from the beginning of this manuscript is actually mentioned by the authors themselves towards the end: the so-called "Winner's Curse". This is the idea, articulated by Andrew Gelman among others, that if an effect is "significant" it is also likely to be an overestimate. In low-powered studies, the estimated effect size can even be opposite in direction to the true effect, yet the p-value is still a low number.  Many ecologists (in my experience) believe that a low p-value indicates two things at the same time: first, that the study "worked"; second, that the effect is real. Neither is necessarily true, and neither can be reliably inferred from the p-value alone. I'm sure the authors are aware of these points, and i think it would be useful to expand on them a bit more. In brief, while I really like the idea of "seeing with clarity", i would appreciate some discussion of how to tell if one has clearly seen a mere mirage._

This is an important concern. We have expanded our discussion of this point, while trying not to stray too far from our main focus.

_I really enjoyed this manuscript and i think ecologists need to read it. I don't want to be negative, but I can't resist a skeptical question: is there any evidence that people will change their behaviours? Misinterpretation of p-values is but one of the many statistical and philosophical errors that pervade science. These will persist until incentive structures change. Until they do, people will continue to abuse their data, their analyses, and their interpretations._

Thanks for your kind words. We agree that this article cannot by itself overcome deeply embedded bad practice, but we hope it will be part of a larger change in a good direction.

### 2019-01-07 doc.tex
BMB general messing around

\noindent In most biological studies, the null hypothesis is known or believed to not be strictly true. Even in cases where the null hypothesis is sensible (e.g., particle physics, \cite{Staley2017}), NHST does not provide evidence that a difference is exactly zero. This being the case, it is worth asking how NHST has survived ``if it is as idiotic as \ldots long believed'' \citet[cited in \cite{Kramer2011}]{ZiliakandMcCloskey2008}.

The value of NHST is that asking whether we can reject the null hypothesis is a proxy for asking whether we see clearly \emph{how} our data differs from it. For example, in a t-test, we are nominally asking whether we can see a \emph{difference} between two means, but the scientific question is whether we are confident which of the two means is \emph{larger}; similarly, tests for whether two values are correlated are a proxy for whether we are confident about the \emph{sign} of the correlation coefficient \citep{robinson2001past}. In other cases (e.g., a one-way ANOVA), it may not be simple to describe the difference we see, but NHST is still a reasonable, widely accepted way to evaluate whether an effect has been seen clearly.

The ``idiocy'', if any, comes in the interpretive step. A statistical fact (``we have \emph{seen} a difference between the groups'', which should immediately prompt the question ``what have you learned \emph{about} that difference?'') is interpreted as a scientific fact (``there \emph{is} a `significant' difference between the groups), which is often seen as an end in itself: ``we showed that the groups differ''.

@@

\noindent We often see sentences like, ``X et al. showed that there is no significant effect of Y on Z'' with the implication that this effect can now be assumed to be absent (or unimportant). In fact, the sentence is erroneous even before we get to the implication: significance tests provide information about \emph{a data set} -- that is, about a study, not about the study system \citep{HoenigandHeisey2001}. Indeed, a very small effect can lead to $p < 0.05$, when data is abundant (or noise is small); or a very large one can lead to $p > 0.05$ when the sample is small or noisy.

@@

Descriptions of statistical results using the language of clarity should begin with reference to the effect. For example, ``The difference between the control and treatment group was not statistically clear.'' Table~\ref{quotetab} shows published examples of statements that misinterpret \pvals in three different ways and demonstrates how to rephrase them in the language of clarity. We have attempted to do this thoughtfully, and therefore the language on the right differs from the language on the left in many ways. The point is not that changing ``significance'' to ``clarity'' is a substantial improvement by itself, but that it can prompt rethinking and reinterpretation. We also hope that, by drawing attention to effects, the language of clarity will encourage more reporting of effect sizes and confidence intervals.

@@

The suggested change in language should help focus attention on improved statistical practice, but does not guarantee it. In particular, there is a danger that ``statistical clarity'' will eventually come to be seen as end to itself, the way that ``significance'' is now. The idea of ``statistical clarity'' will work best if it remains linked to the principles of attaching clarity to the study, not the system, and of focusing on effects.

The use of statistical clarity could be confounded with other meanings of the word clarity. We have found that scientists' understanding of our proposed use of statistical clarity is enhanced by explicitly connecting clarity statments to statements about P values.

ADD a bit about winners' curses, plausibility and the like here (and take some of it out of the Conclusions. Should we say specifically that we are agnostic about (supportive of?) the idea that NHST and/or frequentist stats are overused?

@@

We agree with \citet{Cohen1994} and others \citep{Goodman1999, ZiliakandMcCloskey2008, WassersteinandLazar2016}, that scientific communication and understanding will be improved by a shift away from \pvals to effect sizes and confidence intervals. We argue that the use of ``statistical clarity'' reinforces the need for confidence intervals and effect sizes by making it clearer that bald statements about \pvals are insufficient. The statement ``The difference between our control and treatment groups was not statistically clear ($p = 0.30$)'' is noticeably incomplete; an effect size and confidence interval are required to complete the story.

Improving language will not by itself solve all of the known problems with current statistical practice. We echo previous statements in favor of ``neglected factors'' (prior and related evidence, plausibility of mechanism, study design and data quality, real world benefits, novelty and other factors) \citep{McShaneetal.2017} and reporting of \emph{a priori} analysis of statistical power to avoid emphasis on implausibly large effects given low statistical power \citep[the ``winner's curse''][]{GelmanandCarlin2014, SzucsandIoannidis2017, Bernardietal.2017}. Additionally, we support the idea of writing a statistical journal that chronicles all steps in the analytical process \citep{Kassetal.2016}, and clearly delineating the boundary between inferences based on \emph{a priori} hypotheses and discoveries from \emph{post hoc} data exploration.

>>>>

\noindent In most biological studies, the null hypothesis is known \emph{a priori} to be false. Even in cases where the null hypothesis is sensible (e.g., particle physics, \cite{Staley2017}), NHST does not provide evidence that a difference is exactly zero. This being the case, it is worth asking how NHST has survived ``if it is as idiotic as \ldots long believed'' \citet[cited in \cite{Kramer2011}]{ZiliakandMcCloskey2008}.

The value of NHST is that asking whether we can reject the null hypothesis is a proxy for asking whether we see clearly \emph{how} our data differs from it. For example, in a t-test, we are nominally asking whether we can see a \emph{difference} between two means, but the scientific question is whether we are confident which of the two means is \emph{larger} (and whether the size of the difference is biologically interesting); similarly, tests for whether two values are correlated are a proxy for whether we are confident about the \emph{sign} of the correlation coefficient \citep{robinson2001past}. In other cases (e.g., a one-way ANOVA), it may not be simple to describe the difference we see, but NHST is still a reasonable, widely accepted way to evaluate whether an effect has been seen clearly.

The ``idiocy'', if any, comes in the interpretive step. A statistical fact (``we have \emph{seen} a difference between the groups'', which should immediately prompt the question ``what have you learned \emph{about} that difference?'') is interpreted as a scientific fact (``there \emph{is} a `significant' difference between the groups''), which is often seen as an end in itself: ``we showed that the groups differ''.

@@

\noindent Researchers often write sentences like, ``X et al. showed that there is no significant effect of Y on Z'' with the implication that this effect can now be assumed to be absent (or unimportant). In fact, the sentence is erroneous even before we get to the implication: significance tests provide information about \emph{a data set} -- that is, about a study, not about the study system \citep{HoenigandHeisey2001}. Indeed, a very small effect can lead to $p < 0.05$, when data is abundant (or noise is small); or a very large one can lead to $p > 0.05$ when the sample is small or noisy.

@@

Descriptions of statistical results using the language of clarity should begin with reference to the effect. For example, ``The difference between the control and treatment group was not statistically clear.'' Table~\ref{quotetab} shows published examples of statements that misinterpret \pvals in three different ways and demonstrates how to rephrase them in the language of clarity. We have attempted to do this thoughtfully, and therefore the language on the right differs from the language on the left by more than a simple substitution of ``significance'' to ``clarity''. We do not claim that executing a search-and-replace operation will automatically improve statistical practice; rather, we think it can prompt rethinking and reinterpretation. We also hope that, by drawing attention to effects, the language of clarity will encourage more reporting of effect sizes and confidence intervals.

@@

Changing from ``significance'' to ``clarity'' should help researchers improve their statistical practice, but of course it cannot solve all of our problems with statistics. \bmb{repetitive with \P 3 of Conclusion?}The idea of ``statistical clarity'' will work best if it remains linked to the principles of attaching clarity to the study, rather than the system, and of focusing on effects and confidence intervals. If widely adopted, ``statistical clarity'' could eventually come to be seen as an end in itself, the way that ``significance'' is now; similarly, ``statistical clarity'' could be confounded with other meanings of  ``clarity'' as ``significance'' is now. We hope not, but in any case we feel that the unthinking use of ``clarity'' would be (marginally) better than the current unthinking use of ``significance''. We have found that scientists' understanding of our proposed use of statistical clarity is enhanced by explicitly connecting clarity statements to statements about P values.

@@

We agree with \citet{Cohen1994} and others \citep{Goodman1999, ZiliakandMcCloskey2008, WassersteinandLazar2016}, that scientific communication and understanding will be improved by a shift away from \pvals to effect sizes and confidence intervals. The use of ``statistical clarity'' should reinforce the need for confidence intervals and effect sizes by making bald statements about \pvals more obviously insufficient. The statement ``The difference between our control and treatment groups was not statistically clear ($p = 0.30$)'' is noticeably incomplete; an effect size and confidence interval are required to complete the story.

Improving language will not by itself solve all of the known problems with current statistical practice. We echo previous statements in favor of ``neglected factors'' (prior and related evidence, plausibility of mechanisms, study design and data quality, real world benefits, novelty, etc.) \citep{McShaneetal.2017} and reporting of \emph{a priori} analysis of statistical power to avoid emphasis on implausibly large effects given low statistical power \citep[the ``winner's curse''][]{GelmanandCarlin2014, SzucsandIoannidis2017, Bernardietal.2017}. Additionally, we support the writing of statistical journals that chronicle all of the steps in the analytical process \citep{Kassetal.2016}, and clearly delineating the boundary between inferences based on \emph{a priori} hypotheses and discoveries from \emph{post hoc} data exploration. These procedures help to avoid the ``garden of forking paths'' by which cryptic multiple testing amplifies noise to make it look like a signal of biologically interesting processes \citep{gelman_statistical_2014}).
